---
title: "Example survey analysis"
author: "First_name Last_name"
date: "today"
echo: true
warning: false
message: false
editor: source
format: 
  html:
    code-fold: true
    toc: true
    toc-location: left
    self-contained: true

---

### Load libraries

```{r}
#| label: load_libraries

library(tidyverse)
library(ggtext)
library(gt)
library(patchwork)
library(data.table)

# devtools::install_local('~/Documents/Nature/Programming/surveyR_0.1.0.tgz')
library(surveyR)

```
### Load data

```{r}
#| label: load_data
#| echo: false

mydata <- read_rds(file = 'cleaned_survey_data.rds') |> 
  janitor::clean_names()

```

### Column chunk function
```{r}
# returns a list of columns that match a text string
# inputs: dataset and text string
# outputs: a list of column indices, and a printout of column names
### ADAPTED FROM 4_3_WOS_ANALYSIS.RMD
column_chunk <- function(mytable, text_string, print_colnames = FALSE) {
  questionIDs <- grep(text_string, names(mytable))
  # print relevant column names to check whether last item is a free text field
  if (print_colnames) print(names(mytable[questionIDs]))

  questionIDs
}

```


### Q3
What is your **overall opinion** of the science advice system and practices in your country? 

```{r}
#| label: q3_overall 

full_q <- "Q3: What is your overall opinion of the science advice system and practices in your country?"

levels <- c('Very good (consistent practices and a coherent system)',
            'Good (good practices within many areas of government)',
            'Patchy (science advice practiced inconsistently in government)',
            'Poor (practices are inconsistent and/or of low quality)',
            'Very poor (very little science advice practiced or no system exists)',
            'No opinion/Don’t know')

q3 <- surveyR::process_radio_button_q(mydata, 'what_is_your_overall_opinion_of_the_science_advice_system_and_practices_in_your_country', levels)
footnote <- md(glue::glue("_n = {sum(q3$count)} responses_"))

q3 |> 
  gt::gt(rowname_col = 'option') |> 
  gt::tab_header(title = full_q) |> 
  gt::tab_footnote(footnote) |> 
  gt::fmt_percent('pct')
```

```{r}
#| label: q3_overall_graph

surveyR::graph_responses(q3, full_q) +
  theme(aspect.ratio = 1/10) +
  paletteer::scale_fill_paletteer_d('colorblindr::OkabeIto',
                                    labels = strsplit(levels, "\\(.*\\)$") |>
                                      stringr::str_trim())

```

##### By group

```{r}
#| label: q3_by_group

q3_grp <- surveyR::process_radio_button_q_by_grp(mydata, 'analysis_group',  'what_is_your_overall_opinion_of_the_science_advice_system_and_practices_in_your_country', levels)

footnote <- md(glue::glue("_n = {sum(q3_grp$count, na.rm = T)} responses_"))
q3_grp |> 
  gt::gt(groupname_col = 'analysis_group', rowname_col = 'option') |> 
  gt::tab_header(title = full_q) |> 
  gt::tab_footnote(footnote) |> 
  gt::fmt_percent(columns = 'pct') |> 
  gt::sub_missing() |> 
  gt::summary_rows(fns = list(id = 'sum', label = md("**TOTAL**"), fn = 'sum'))
```

```{r}
#| label: q3_by_group_graph

surveyR::graph_responses_by_grp(q3_grp, 'analysis_group', full_q) +
  theme(aspect.ratio = 3/10) +
  paletteer::scale_fill_paletteer_d('colorblindr::OkabeIto', 
                                    labels = strsplit(levels, "\\(.*\\)$") |>
                                      stringr::str_trim())  

```

##### By World Bank income group

```{r}
#| label: q3_by_wb

q3_wb <- surveyR::process_radio_button_q_by_grp(mydata, 'income_group', 'what_is_your_overall_opinion_of_the_science_advice_system_and_practices_in_your_country', levels)

footnote <- md(glue::glue("_n = {sum(q3_wb$count, na.rm = T)} responses_"))

q3_wb |> 
  filter(!is.na(income_group)) |> 
  gt::gt(groupname_col = 'income_group', rowname_col = 'option') |> 
  gt::tab_header(title = full_q) |> 
  gt::tab_footnote(footnote) |> 
  gt::fmt_percent(columns = 'pct') |> 
  gt::sub_missing() |> 
  gt::summary_rows(fns = list(id = 'sum', label = md("**TOTAL**"), fn = 'sum'))
```

```{r}
#| label: q3_by_wb_graph

surveyR::graph_responses_by_grp(q3_wb, 'income_group', full_q) +
  theme(aspect.ratio = 3/10) +
  paletteer::scale_fill_paletteer_d('colorblindr::OkabeIto', 
                                    labels = strsplit(levels, "\\(.*\\)$") |>
                                      stringr::str_trim())  

```

### Q6
Please indicate how strongly you agree or disagree with the following statements about **obstacles to successful science advice to government in your country**. 

```{r}
#| label: q6_overall

levels <- c('Strongly agree', 'Agree', 'Neither agree nor disagree', 'Disagree', 'Strongly disagree')
full_q <- 'Q6: Please indicate how strongly you agree or disagree with the following statements about obstacles to successful science advice to government in your country.'
tmp <- surveyR::process_multiq_radio_button_q(mydata, 'please_indicate_how_strongly_you_agree_or_disagree_with_the_following_statements_about_obstacles_to_successful_science_advice', levels)
tmp2 <- as_tibble(data.table::rbindlist(tmp, idcol = TRUE))

t <- tmp2 |>
  mutate(option = factor(option, levels = levels, ordered = TRUE)) |>
  arrange(option) |>
  filter(option != 'NA' & !is.na(option)) |>
  group_by(.id) |> 
  mutate(pct = count/sum(count)) |> 
  gt(groupname_col = '.id', rowname_col = 'option') |>
  gt::summary_rows(fns = list(id = 'sum', label = md('**TOTAL RESPONSES**')) ~ sum(., na.rm = T)) |> 
  gt::tab_header(full_q) |>
  gt::sub_missing(missing_text = '--') |>
  gt::fmt_percent('pct')

t
```

```{r}
#| label: q6_overall_graph
#| fig-width: 10
#| fig-height: 13

l <- list() # list to hold the graphs
v <- NA # empty vector to hold the positions of graphs
for (i in 1:length(tmp)) {
  l[[i]] <- surveyR::graph_responses(tmp[[i]], names(tmp[i]))
  v[i] <- l[[i]]$data |> 
    # mutate(fxn = value/count) |> 
    filter(option == 'Strongly agree') |> 
    # purrr::pluck('fxn')
    purrr::pluck('pct')
}
# number the vector
names(v) <- 1:length(tmp)
# pos is a list of indices, sorted by level of 'strongly agree'
pos <- as.integer(names(sort(v, decreasing = TRUE)))

final <-  l[[pos[1]]] 
for (i in 2:length(l)) {
  final <- final / l[[pos[i]]]  
}
final <- final +
  patchwork::plot_annotation(title = full_q) +
  patchwork::plot_layout(guides = 'collect') &
  theme(legend.position = 'bottom')
final

```
##### By group

```{r}
#| label: q6_by_group
#| results: asis

tmp_grp <- surveyR::process_multiq_radio_button_q_by_grp(mydata, 'analysis_group',  'please_indicate_how_strongly_you_agree_or_disagree_with_the_following_statements_about_obstacles_to_successful_science_advice', levels)

l <- list()
for (i in 1:length(tmp_grp)) {
  mysubtitle <- names(tmp_grp[i])
  footnote = md(glue::glue("_n = {sum(tmp_grp[[i]]$count, na.rm = TRUE)} responses_"))
  
  l[[i]] <- tmp_grp[[i]] |>
    filter(!is.na(analysis_group)) |> 
    gt::gt(groupname_col = 'analysis_group', rowname_col = 'option') |> 
    gt::tab_header(title = full_q, subtitle = mysubtitle) |> 
    gt::tab_footnote(footnote) |> 
    gt::fmt_percent(columns = 'pct') |> 
    gt::sub_missing() |> 
    gt::summary_rows(fns = list(id = 'sum', label = md("**TOTAL**"), fn = 'sum'))
}

l

```

```{r}
#| label: q6_by_group_graph
#| fig-width: 10
#| fig-height: 15

l <- list() # list to hold the graphs
for (i in 1:length(tmp_grp)) {
  l[[i]] <- surveyR::graph_responses_by_grp(tmp_grp[[i]], 'analysis_group', names(tmp_grp[i]))
}

final <-  l[[1]] 
for (i in 2:length(l)) {
  final <- final / l[[i]]  
}
final <- final +
  patchwork::plot_annotation(title = full_q) +
  patchwork::plot_layout(guides = 'collect') &
  theme(legend.position = 'bottom')
final

```
##### By World Bank income group

```{r}
#| label: q6_by_wb
#| results: asis

tmp_wb <- surveyR::process_multiq_radio_button_q_by_grp(mydata, 'income_group',  'please_indicate_how_strongly_you_agree_or_disagree_with_the_following_statements_about_obstacles_to_successful_science_advice', levels)

l <- list()
for (i in 1:length(tmp_wb)) {
  mysubtitle <- names(tmp_wb[i])
  footnote = md(glue::glue("_n = {sum(tmp_grp[[i]]$count, na.rm = TRUE)} responses_"))
  
  l[[i]] <- tmp_wb[[i]] |>
    filter(!is.na(income_group)) |> 
    gt::gt(groupname_col = 'income_group', rowname_col = 'option') |> 
    gt::tab_header(title = full_q, subtitle = mysubtitle) |> 
    gt::tab_footnote(footnote) |> 
    gt::fmt_percent(columns = 'pct') |> 
    gt::sub_missing() |> 
    gt::summary_rows(fns = list(id = 'sum', label = md("**TOTAL**"), fn = 'sum'))
}

l

```

```{r}
#| label: q6_by_wb_graph
#| fig-width: 10
#| fig-height: 15

l <- list() # list to hold the graphs
for (i in 1:length(tmp_wb)) {
  l[[i]] <- surveyR::graph_responses_by_grp(tmp_wb[[i]], 'income_group', names(tmp_wb[i]))
}

final <-  l[[1]] 
for (i in 2:length(l)) {
  final <- final / l[[i]]  
}
final <- final +
  patchwork::plot_annotation(title = full_q) +
  patchwork::plot_layout(guides = 'collect') &
  theme(legend.position = 'bottom')
final

```

### Q9
Looking internationally, **which country** do you think is **particularly successful** at ensuring science is factored into government policies and decisions i.e. they have an enviable science advice system?

```{r}
#| label: q9_overall

# this list of countries is more or less comprehensive given the answers to q9 and q10:
re_usa <- '\\bUSA?\\b|United States( of America)?|\\bEEUU\\b|U\\.S\\.(A\\.)?'
terms <- c(
  re_usa,
  'United Kingdom|\\bUK\\b|Great Britain|England|Scotland|U\\.K\\.',
  # either " EU" or "^EU" but not "EEUU", eg
  '\\bEU\\b|European (Union|Commission)',
  'Afghanistan',
  'Albania',
  'Algeria',
  'Angola',
  'Argentina',
  'Australia',
  'Bangladesh',
  'Belarus',
  'Belgium',
  'Botswana',
  'Bra[zs]il',
  'Burundi',
  'Cambodia',
  'Camero[ou]n',
  'Canad[az]',
  'Centrafrique|Central African Republic',
  'Chad',
  'Chil[ei]',
  'China',
  'Colombia',
  'Congo',
  'Costa Rica',
  'Cuba',
  'Ecuador',
  'El Salvador',
  'Eritrea',
  'Ethiopia',
  'Franc[e|ia]',
  '[G|H]ermany|Alemania|Deutchland',
  'Ghana',
  'Greece',
  'Guatemala',
  'Haiti',
  'Honduras',
  'Hungary',
  'India',
  'Indonesia',
  'Iran',
  'Ireland',
  'Israel',
  'Italy',
  'Japan',
  'Kenya',
  'Latvia',
  'Malay?sia',
  'Mexico',
  'Myanmar',
  'Netherlands?',
  'N(ew|ueva)[ -]Zea?landa?|NZ',
  'Nicaragua',
  'Nigeria',
  'North Korea',
  'Pakistan',
  'Peru',
  'Philippines',
  'Poland',
  'Portugal',
  'Romania',
  'Russia',
  'Rwanda',
  'Samoa',
  'Scandinavia|Finland|Norway|Sweden|D[ei]na?mark|Nordic',
  'Singap[ou]re',
  'Somalia',
  'South Africa',
  'South Korea|(?<!North )Korea', # (?<!...) allows Korea but not North Korea
  '(South )?Sudan',
  'Spain',
  'Sri Lanka',
  'Switzerland|Suiza',
  'Taiwan',
  'Togo',
  'Turkey',
  'Venezuela',
  'Vietnam',
  'Zimbabwe'
)

question <- "Q9: Looking internationally, which country do you think is particularly successful at ensuring science is factored into government policies and decisions?"

surveyR::process_free_text_q(mydata, terms = terms, question = 'looking_internationally_which_country_if_any_do_you_think_is_particularly_successful_at_ensuring_science_is_factored_into_government_policies_and_decisions_i_e_they_have_an_enviable_science_advice_system')

```


### Q10
Looking internationally, **which country** do you think is **particularly unsuccessful** at ensuring science is factored into government policies and decisions? 

```{r}
#| label: q10_overall

question <- "Q10: Looking internationally, which country do you think is particularly unsuccessful at ensuring science is factored into government policies and decisions?"

surveyR::process_free_text_q(mydata, terms = terms, question = 'looking_internationally_which_country_if_any_do_you_think_is_particularly_unsuccessful_at_ensuring_science_is_factored_into_government_policies_and_decisions')

```

### Q28
Which of these best describes the organization(s) you work for?

```{r}
#| label: q28_overall

tmp <- surveyR::process_check_all_that_apply_q(mydata, 'which_of_these_best_describes_the_organization_s_you_work_for') |> 
  filter(option != "other write in 78") 

# get total number of respondents (people could select >1 answer)
index <- column_chunk(mydata, 'which_of_these_best_describes_the_organization_s_you_work_for')
response_count <- mydata |> 
  select(all_of(index)) |> 
  rowwise() %>% 
  mutate(any_response = any(!is.na(across(everything())))) |> 
  # if all responses are NA, all will be FALSE and any_response will be 0, logical FALSE
  # convert to logical rather than int to count any response at all
  purrr::pluck('any_response') |> 
  sum()

footnote <- md(glue::glue("_n = {response_count} respondents_"))

tmp |> 
  arrange(desc(count)) |> 
  gt::gt(rowname_col = 'option') |> 
  gt::tab_header("Q28: Which of these best describes the organization(s) you work for?") |> 
  gt::tab_footnote(footnote)
```
